Constraints on the mass and occupation fraction of dark matter halos hosting Lyman-alpha emitters at $z\sim 3$

### 2013-07-12 paper/ms.tex
Paulina comments partially added

This complexity makes the use of monte-carlo simulations for the
radiative transfer a required tool to obtain physically sound results,
although the degeneracy in the physical parameters involved in the
problem makes it difficult to achieve a robust consensus on what is
the theoretical expected value for the \ly escape fraction
at high redshift.

@@

Our method is based on the comparison of observatios and mock
catalogs on the surface number density and the angular correlation
function.  In the next subsections we describe in detail the four key
elements of this workflow. First, we present the observations we take
as a benchmark. Second, the N-body simulation and the halo catalogs we
use. Third, the simplified model that allows us to translated halo
catalogs into mock LAE observations. Fourth, the statistics we use to
compare observational results against our theoretical predictions.

@@

One of the main observational benchmarks we use in this paper is the LAE number
density information at $z=3.1$ obtained by the panoramic narrow-band
survey presented by \cite{Yamada2012} from a survey  conducted with
the Subaru 8.2m telescope and the Subaru Prime Focus Camera, which has
a field of view covering $34\times 27$ arcmin, corresponding to a
comoving scale of $46\times35$ Mpc $h^{-1}$ at $z=3.09$. The narrow
band filter is centered at $4977$ \AA with  $77$\AA width,
corresponding to the redshift range $z=3.062-3.125$ and $41$ Mpc
$h^{-1}$ comoving scale for the detection of the Lyman-$\alpha$ line
centered at $z=3.09$. The authors report a total  $2161$ LAEs with an
observed equivalent width larger than $190$\ over a total survey area
of $2.42$ deg$^{2}$, this corresponds to  average surface number
density of $0.20\pm 0.01$ arcmin$^{-2}$.

The survey covered four independent fields. The first is the SSA22
field of $1.38$ deg$^2$ with $1394$ detected LAEs, this field has been
known to harbor a region with a large density excess of galaxies. The
second observed region is composed by the fields Subaru/{\it
  XMM-Newton} Deep Survey (SXDS)-North, -Center and -South, with a 
total of $0.58$ deg$^2$ and $386$ LAEs. The third and fourth fields
are the Subaru Deep Field (SDF) with $0.22$ deg$^2$ and $196$ LAEs,
and the field around the Great Observatory Optical Deep Survey North
(GOODS-N) with $0.24$ deg$^2$ and $185$ LAEs.

There is abundant observational work done on LAEs at redshift $z=3.1$
\citep{Kudritzki2000,Matsuda2005,Gawiser2007,Nilsson2007,Ouchi2008}.
However, we decide to focus on the data from \cite{Yamada2012} because
it has the largest covered area with homogeneous instrumentation
conditions (telescope, narrow band filter), data reduction pipeline
and conditions to construct the LAE catalog. This ensures that the
number density variations among fields are due only to
astrophysical reasons and not different observational conditions or
criteria to construct the catalogs.

@@

We use halo catalogs constructed with a Friend-of-Friends (FOF)
algorithm with a linking length of 0.17 times the inter-particle
distance. The minimum halo mass in the models we construct in this \documentname
correspond to groups of $\sim 75$ particles. The catalogs were
obtained from the publicly available Multidark database \footnote{{\tt
    http://www.multidark.org/MultiDark/}} \citep{2011arXiv1109.0003R}.
We focus our work on halos more massive than $1\times 10^{10}$\hMsun
that are resolved with at least $70$ particles.

@@

In our model a dark matter halo can only host one or zero LAE. 
There are three parameters in the model that decide whether a halo can host a
galaxy or not: the lower and upper bounds for the mass range $M_{\rm min}<
M_{\rm h} < M_{\rm max}$ where LAEs reside and the fraction $f_{\rm
  occ}$ of such halos that host a \emph{detectable LAE}. The reader
must keep in mind that the physical interpreation of the occupation
fraction $f_{\rm occ}$ convolves two things: the presence of a star
forming galaxy in a halos and its detectability as a LAE. We do not assign a
luminosity to each LAE. We are primarily interested in
constraining the halo mass range hosting detectable LAEs under the
conditions defined by \cite{Yamada2012}. In what follows we will describe by
the letter ${\mathcal M}$ a model defined by an specific choice of the
three scalar parameters $M_{\rm min}$, $M_{\rm  max}$ y $f_{\rm occ}$.

@@

The second benchmark is the angular correlation function (ACF). The
measurements presented in \citep{Yamada2012} do not report an ACF
measurement. Instead we use the results reported by
\citep{Hayashino2004} on the LAEs observed in the densest field of
SSA22. It is important to keep in mind that there are some differences
between this work and \citep{Yamada2012}. First, the color selection by
\cite{Yamada2012} is less stringent compared to the one by
\cite{Hayashino2004}. Also the EW threshold is different,
\cite{Hayashino2004} uses a cut around $154$\AA \ instead of $190$\AA
\ used by \citep{Yamada2012}

@@

The second region corresponds to an intermediate range for the minimum
mass $10.5<\log_{10}M_{\rm min}<11.5$ where regardless of the value of
the maximum mass $M_{\rm max}$ it is possible to tune the occupation
fraction $f_{\rm occ}$ to bring some of the mock observations into
good agreement with observations. In this region in parameter space
one can thus find two extreme kinds of models.  One kind where the
mass interval is very narrow with sizes smaller than $<0.3$ dex (a
factor of two in mass) and others where the mass interval is 
extended, larger than $1.0$ dex, going up to the maximum halo mass
present in the simulation at that redshift.

@@

Without any additional information our method allows us to infer that
most of the successful models are found in the second and third region of
parameter space where. This result was already expected from halo
abundance calculations shown in Figure \ref{fig:halos}. In the next
section we reduce the size of this region by imposing tighter
constraints on the KS tests and consistency with the angular
correlation function.

@@

Figure \ref{fig:restriction_mock} presents the locii of these models
in the parameter space $M_{\rm min}-M_{\rm max}$ and $M_{\rm
  min}-f_{\rm occ}$. The results are very similar between the {\texttt
  match} and {\texttt random} methods. With this constraints the
number of consistent models with  $\log_{10}M_{\rm min}< 11.5$ are
greatly reduced. This corresponds to the regions in the parameter
space in Figure \ref{fig:landscape} that already had a low number of
consistent mock surveys. On the other hand, from the right panel in
Figure \ref{fig:restriction_mock} one can see that there is not a
strong selection effect on the occupation fraction $f_{\rm occ}$.

@@

In order to better illustrate the results from this test we
show in the right panel of Figure \ref{figure:correlation_parameters} 
the plane $\theta_{0}$-$M_{\rm min}$ for all models. There we divide
the models into two disjoint sets: those with $\log_{10} M_{\rm
    max}<12$ (blue dots) and $\log_{10} M_{\rm
    max}>12$ (green dots). The colored rectangle includes
the parameter region which is consistent with the observational
constraint in $\theta_{0}$. With this restriction we find that most
models with $\log_{10} M_{\rm max}>12$ and  $\log_{10} M_{\rm min}>11.1$ 
%can be safely ruled out.

@@

Matching the galaxy surface number density statistics sets the median
mass of all successful models in the range $(10^{10}-10^{12})$
\hMsun. Including more strict criteria on the number of mock surveys
that must be consistent with those observations and additional
information from the angular correlation function greatly reduces the
number of possible models. We end up with $\sim 50$ models out of the
initial 9000 possible combination of parameters.

@@

One possible interpretation can be made in terms of a decreasing escape
fraction of \ly radiation in massive systems. There are detailed models for
radiative transfer that support the idea that massive galaxies with
higher metallicities have larger dust contents and less concentrated
ISM than lower mass systems, which due to the resonant nature of the
\ly line are enough to produce high absorption of \ly photons but not of continuum or
other non-resonant lines \citep{ForeroRomero2011,Laursen2009}.

@@

The preference for narrow $\Delta M$ ranges to hosting LAEs together
with the minor presence of reasonable models with $M_{\rm max}>
10^{12}$\hMsun \  shows that our models support theoretical insights
where the most massive systems are not bright \ly sources
\citep{ForeroRomero2012}.

>>>>

This complexity can be tackeld with monte-carlo simulations for the
radiative transfer a required tool to obtain physically sound results,
although the degeneracy in the physical parameters involved in the
problem makes it difficult to achieve a robust consensus on what is
the theoretical expected value for the \ly escape fraction
at high redshift.

@@

Our method is based on the comparison of observatios and mock
catalogs on the surface number density and the angular correlation
function of LAEs.  In the next subsections we describe in detail the four key
elements of this workflow. First, we present the observations we take
as a benchmark. Second, the N-body simulation and the halo catalogs we
use. Third, the simplified model that allows us to translated halo
catalogs into mock LAE observations. Fourth, the statistics we use to
compare observational results against our theoretical predictions.

@@

One of the main observational benchmarks we use in this paper is the LAE number
density information at $z=3.1$ obtained by the panoramic narrow-band
survey presented by \cite{Yamada2012} from a survey
conducted with the Subaru 8.2m telescope and the Subaru Prime Focus Camera,
which has a field of view covering $34\times 27$ arcmin, corresponding to a
comoving scale of $46\times35$ Mpc $h^{-1}$ at $z=3.09$.  The narrow
band filter is centered at $4977$ \AA with  $77$\AA width,
corresponding to the redshift range $z=3.062-3.125$ and $41$ Mpc
$h^{-1}$ comoving scale for the detection of the Lyman-$\alpha$ line
centered at $z=3.09$. The authors reported a
total  $2161$  LAEs with an observed equivalent width larger than $190$\AA 
over a total survey area of $2.42$ deg$^{2}$ that includes 12 subfields, 
this corresponds to average surface number density of $0.20\pm 0.01$ arcmin$^{-2}$.

The survey covered four independent fields. The first is the SSA22
field of $1.38$ deg$^2$ with $1394$ detected LAEs (7 subfields), this field has been
known to harbor a region with a large density excess of galaxies. The
second observed region is composed by the fields Subaru/{\it
  XMM-Newton} Deep Survey (SXDS)-North, -Center and -South, with a 
total of $0.58$ deg$^2$ and $386$ LAEs (3 subfields). The third and fourth fields
are the Subaru Deep Field (SDF) with $0.22$ deg$^2$ and $196$ LAEs (1 subfield),
and the field around the Great Observatory Optical Deep Survey North
(GOODS-N) with $0.24$ deg$^2$ and $185$ LAEs (1 subfield).

There is abundant observational work done on LAEs at redshift $z=3.1$
\citep{Kudritzki2000,Matsuda2005,Gawiser2007,Nilsson2007,Ouchi2008}.
However, we decide to focus on the data from \cite{Yamada2012} because
it has the largest covered area with homogeneous instrumentation
conditions (telescope, narrow band filter), data reduction pipeline
and conditions to construct the LAE catalog. This ensures that the
number density variations among fields are \emph{not} due to different observational
conditions or criteria to construct the catalogs.

The second constraint is the angular correlation function (ACF). The
measurements presented in \citep{Yamada2012} do not report an ACF
measurement. Instead we use the results reported by
\citep{Hayashino2004} on the LAEs observed in the densest field of
SSA22. It is important to keep in mind that there are some differences
between this work and \citep{Yamada2012}. First, the color selection by
\cite{Yamada2012} is less stringent compared to the one by
\cite{Hayashino2004}. Also the EW threshold is different,
\cite{Hayashino2004} uses a cut around $154$\AA \ instead of $190$\AA
\ used by \citep{Yamada2012}

@@

We use halo catalogs constructed with a Friend-of-Friends (FOF)
algorithm with a linking length of 0.17 times the inter-particle
distance. The minimum halo mass in the models we construct in this \documentname
correspond to groups of $\sim 75$ particles. The catalogs were
obtained from the publicly available Multidark database \footnote{{\tt
    http://www.multidark.org/MultiDark/}} \citep{2011arXiv1109.0003R}.
We focus our work on halos more massive than $1\times 10^{10}$\hMsun
that are resolved with at least $70$ particles (REFF????)

@@

In our model a dark matter halo can only host one or zero LAE. 
There are three parameters in the model that decide whether a halo can host a
galaxy or not: the lower and upper bounds for the mass range $M_{\rm min}<
M_{\rm h} < M_{\rm max}$ where LAEs reside and the fraction $f_{\rm
  occ}$ of such halos that host a \emph{detectable LAE}. The reader
must keep in mind that the physical interpreation of the occupation
fraction $f_{\rm occ}$ convolves two things: the presence of a star
forming galaxy in a halo and its detectability as a LAE. We do not assign a
luminosity to each LAE. We are primarily interested in
constraining the halo mass range hosting detectable LAEs under the
conditions defined by \cite{Yamada2012}. In what follows we will describe by
the letter ${\mathcal M}$ a model defined by an specific choice of the
three scalar parameters $M_{\rm min}$, $M_{\rm  max}$ and $f_{\rm occ}$.

@@

@@

The second region corresponds to an intermediate range for the minimum
mass $10.5<\log_{10}M_{\rm min}<11.5$ where regardless of the value of
the maximum mass $M_{\rm max}$ it is possible to tune the occupation
fraction $f_{\rm occ}$ to bring some of the mock observations into
good agreement with observations. In this region in parameter space
one can thus find two extreme kinds of models.  One kind where the
mass interval is very narrow with sizes smaller than $<0.3$ dex (a
factor of two in mass) and others where the mass interval is 
extended (larger than $1.0$ dex) going up to the maximum halo mass
present in the simulation at that redshift (up to $2.5$ dex).

@@

Without any additional information our method allows us to infer that
most of the successful models are found in the second and third region of
parameter space where this result was already expected from halo
abundance calculations shown in Figure \ref{fig:halos}. In the next
section we reduce the size of this region by imposing tighter
constraints on the KS tests and consistency with the angular
correlation function.

@@

Figure \ref{fig:restriction_mock} presents the locii of these models
in the parameter space $M_{\rm min}-M_{\rm max}$ and $M_{\rm
  min}-f_{\rm occ}$. The results are very similar between the {\texttt
  match} and {\texttt random} methods. With this constraints the
number of consistent models with  $\log_{10}M_{\rm min}< 11.5$ are
greatly reduced. This corresponds to the regions in the parameter
space in Figure \ref{fig:landscape} that already had a low number of
consistent mock surveys. On the other hand, from the right panel in
Figure \ref{fig:restriction_mock} one can see that there is not a
strong constraint on the occupation fraction $f_{\rm occ}$.

@@

In order to better illustrate the results from this test we
show in the right panel of Figure \ref{figure:correlation_parameters} 
the plane $\theta_{0}$-$M_{\rm min}$ for all models. There we divide
the models into two disjoint sets: those with $\log_{10} M_{\rm
    max}<12$ (blue dots) and $\log_{10} M_{\rm
    max}>12$ (green dots). The colored rectangle includes
the parameter region which is consistent with the observational
constraint in $\theta_{0}$. With this restriction we find that most
models with $\log_{10} M_{\rm max}>12$ and  $\log_{10} M_{\rm min}>11.1$ 
can be safely ruled out.

@@

Matching the galaxy surface number density statistics to
our mock  data sets the median
mass of all successful models in the range $(10^{10}-10^{12})$
\hMsun. Including more strict criteria on the number of mock surveys
that must be consistent with those observations and additional
information from the angular correlation function greatly reduces the
number of possible models. We end up with $\sim 50$ models out of the
initial 9000 possible combination of parameters.

@@

One possible interpretation can be made in terms of a decreasing escape
fraction of \ly radiation in massive systems. There are detailed models for
radiative transfer that support the idea that massive galaxies with
higher metallicities have larger dust contents and less concentrated
ISM than lower mass systems which, due to the resonant nature of the
\ly line that increase the probability of absortion  of\ly photons, produce
high absorption of \ly photons but not of continuum or
other n-resonant lines \citep{ForeroRomero2011,Laursen2009}.

@@

The preference for narrow $\Delta M$ ranges to hosting LAEs together
with the minor presence of reasonable models with $M_{\rm max}>
10^{12}$\hMsun \  shows that our models support theoretical and observational
insights where the most massive systems are not bright \ly sources
\citep{ForeroRomero2012}, \citep{Shapley2003}
